# Continuous G {#sec-continuous}

```{r setup-cont, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
```

This chapter demonstrates FDID estimation when the baseline factor $G$ is continuous,
using `lnpczupu` (log genealogy book density) as the treatment variable.

::: {.callout-note appearance="simple"}
**Note**: Continuous $G$ currently supports only the linear estimators (`"did"`, `"ols1"`, `"ols2"`).
Methods that require binary group membership — `"ebal"`, `"ipw"`, `"aipw"` — are not available.
Support will be extended in future versions.
:::

## Setup

```{r library-cont}
library(fdid)
data(fdid)
```

```{r data-cont}
mortality$uniqueid <- paste(as.character(mortality$provid),
                            as.character(mortality$countyid), sep = "-")

covar      <- c("avggrain", "nograin", "urban", "dis_bj", "dis_pc",
                "rice", "minority", "edu", "lnpop")
tr_period  <- 1958:1961
ref_period <- 1957
```

Prepare the data using `lnpczupu` as the continuous $G$:

```{r prepare-cont}
s_cont <- fdid_prepare(
  data       = transform(mortality, G = lnpczupu),
  Y_label    = "mortality",
  X_labels   = covar,
  G_label    = "G",
  unit_label = "uniqueid",
  time_label = "year"
)
```

## Estimation

### DID (no covariates)

```{r cont-did}
result_cont_did <- fdid(s_cont, tr_period = tr_period, ref_period = ref_period,
                        method = "did")
summary(result_cont_did)
```

### OLS without interactions

```{r cont-ols1}
result_cont_ols1 <- fdid(s_cont, tr_period = tr_period, ref_period = ref_period,
                         method = "ols1")
summary(result_cont_ols1)
```

### OLS with interactions (OLS$_*$)

```{r cont-ols2}
result_cont_ols2 <- fdid(s_cont, tr_period = tr_period, ref_period = ref_period,
                         method = "ols2")
summary(result_cont_ols2)
```

## Plots

```{r cont-plot-raw}
plot(result_cont_ols2, type = "raw")
```

```{r cont-plot-dynamic}
plot(result_cont_ols2, type = "dynamic")
```
