# Visualization Options {#sec-visualization}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
```

In this chapter, we explore various visualization options available in the `fdid` package.

```{r}
library(fdid)
```

```{r readcsv, include=FALSE}
data(fdid)
```

```{r gen, include=FALSE}
mortality$uniqueid <- paste(as.character(mortality$provid), as.character(mortality$countyid), sep = "-")
mortality$highzupu <- ifelse(mortality$pczupu >= median(mortality$pczupu, na.rm = TRUE), 1, 0)
```

```{r parameters, include=FALSE}
tr_period <- 1958:1961
ref_period <- 1957
treat <- "highzupu"
Y <- "mortality"
time <- "year"
unit <- "uniqueid"
covar <- c("avggrain", "nograin", "urban", "dis_bj", "dis_pc",
           "rice", "minority", "edu", "lnpop")
```

```{r prepare, include=FALSE}
prepared_d <- fdid_prepare(
  data = mortality,
  Y_label = Y,
  X_labels = covar,
  G_label = treat,
  unit_label = unit,
  time_label = time
)
```

```{r ols1, include=FALSE}
fdid_results_ols1 <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  vartype = "robust",
)
```

```{r ols2, include=FALSE}
fdid_results_ols2 <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ols2",
  vartype = "robust",
)
```

```{r did, include=FALSE}
fdid_results_did <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "did",
  vartype = "robust",
)
```

```{r ebal, cache=TRUE, include=FALSE}
fdid_results_ebal <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ebal",
  vartype = "robust",
  target.pop = "1"
)
```

```{r ipw, cache=TRUE, include=FALSE}
fdid_results_ipw <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ipw",
  vartype = "robust",
)
```

```{r aipw, cache=TRUE, include=FALSE}
fdid_results_aipw <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "aipw",
  vartype = "robust",
)
```

```{r boot, cache=TRUE, include=FALSE}
fdid_results_ols1_boot <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  vartype = "bootstrap",
  parallel = TRUE,
  cores = 6,
  nsims = 100
)
```

```{r jackknife, cache=TRUE, include=FALSE}
fdid_results_ols1_jack <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  vartype = "jackknife",
  parallel = TRUE,
  cores = 6,
)
```

```{r ols2_target1, cache=TRUE, include=FALSE}
fdid_results_ols2_target_1 <- fdid(
  s = prepared_d,
  tr_period = tr_period,
  ref_period = ref_period,
  method = "ols2",
  vartype = "robust",
  target.pop = "1"
)
```


## Raw means

-  `ci = FALSE` removes the 95% confidence intervals.

```{r}
plot(fdid_results_ols1, type = "raw", ci = FALSE)
```

-  `connected = TRUE` connects points with lines (default: `FALSE`).

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE)
```

-  `group_labels` renames the two groups.

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE, 
     group_labels = c("Low Social Capital", "High Social Capital"))
```

-   `palette` chooses an RColorBrewer palette (default: "Set2").

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE, 
     group_labels = c("Low Social Capital", "High Social Capital"), palette = "Set1")
```

-   `shade_periods` controls the position of shaded region. Default is `x$tr_period`, i.e. the event window.     
    - By setting `shade_periods = c(1954, 1957)`, the pre-event periods can be set shaded.
    - By setting `shade_periods = NULL`, the user can also remove the shaded region.

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE, 
     group_labels = c("Low Social Capital", "High Social Capital"),
     shade_periods = c(1954, 1957))
```

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE, 
     group_labels = c("Low Social Capital", "High Social Capital"),
     shade_periods = NULL)
```

-   `alpha_shade` controls the transparency of the shaded region.

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE, 
     group_labels = c("Low Social Capital", "High Social Capital"),
     alpha_shade = 0.6)
```

-   The user can also set `ylim`, `xlab`, `ylab`, and `main` (currently, xlim is not supported).

```{r}
plot(fdid_results_ols1, type = "raw", connected = TRUE, 
     group_labels = c("Low Social Capital", "High Social Capital"), 
     ylim = c(0, 80), xlab = "Year", ylab = "Mortality", main = "Raw Mortality")
```

## Dynamic estimates

-   `ci = FALSE` removes the 95% confidence intervals.

```{r}
plot(fdid_results_ols1, type = "dynamic", ci = FALSE)
```

-   `connected = TRUE` connects points with lines (default: `FALSE`).

```{r}
plot(fdid_results_ols1, type = "dynamic", connected = TRUE)
```

-   `shade_periods`, `alpha_shade`, `ylim`, `xlab`, `ylab`, and `main` customize the plot as above.

```{r}
plot(fdid_results_ols1, type = "dynamic", connected = TRUE,
     shade_periods = c(1954, 1957),
     alpha_shade = 0.8, ylim = c(-10, 4),
     ylab = "Dynamic Moderating Effect", xlab = "Year", 
     main = "Dynamic Effect of High Social Capital"
     )
```

-   `xlim` zooms into a subset of years.

```{r}
plot(fdid_results_ols1, type = "dynamic",
     alpha_shade = 0.8, ylim = c(-10, 4), xlim = c(1954, 1961),
     ylab = "Dynamic Moderating Effect", xlab = "Year", 
     main = "Dynamic Effect of High Social Capital w/o Post-Event Periods"
     )
```

## Overlap test 

The user can adjust `palette`, `ylim`, `xlab`, and `main`:

```{r}
plot(fdid_results_aipw, type = "overlap",
     palette = "Set1", ylim = c(-0.1, 0.1),
     xlab = 'Score', main = 'Diagnostic Test for Overlap Assumption'
     )
```

## Comparison of methods

By default, the comparison plot uses a vertical layout (estimates stacked top to bottom). Set `vertical = FALSE` for a horizontal layout.

```{r}
plot(fdid_list(fdid_results_did, fdid_results_ols1, fdid_results_ols2, 
               fdid_results_ebal, fdid_results_aipw), vertical = FALSE
     )
```

Common options:

-   `xlab` and `main` change labels/titles.

-   `ylim` sets the range of the estimates axis.

```{r}
plot(fdid_list(fdid_results_did, fdid_results_ols1, fdid_results_ols2, 
               fdid_results_ebal, fdid_results_aipw),
     ylim = c(-5,5)
     )
```

```{r}
plot(fdid_list(fdid_results_did, fdid_results_ols1, 
               fdid_results_ols2, fdid_results_ebal, fdid_results_aipw),
     xlab = 'Static Estimates',
     main = 'Comparing Different Methods', vertical = FALSE,
     ylim = c(-10,10)
     )
```

The labels reflect `method`, and add `vartype` when it is not the default `"robust"`. Other settings (e.g., `target.pop`) are not currently shown in the labels.

```{r}
plot(fdid_list(fdid_results_ols1, fdid_results_ols1_boot,
               fdid_results_ols1_jack, fdid_results_ols2, fdid_results_ols2_target_1,
               fdid_results_aipw),
     ylim = c(-5,1), vertical = TRUE
     )
```

