# Sensitivity Analysis {#sec-sensitivity}

```{r setup-sens, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
```

After estimating treatment effects with `fdid()`, a natural next step is to
assess how robust the results are to potential unmeasured confounders. We
recommend using the [`sensemakr`](https://carloscinelli.com/sensemakr/) package
(@cinelli2020making) for this purpose.

`sensemakr` quantifies sensitivity in terms of partial $R^2$: how strongly
would an omitted variable need to be associated with both the treatment and the
outcome to explain away the estimated effect. A full methodological discussion
is in Xu, Zhao, and Ding (2026, Appendix A3).

## Setup

```{r library-sens}
library(fdid)
library(sensemakr)
data(fdid)
```

Recreate the prepared data and define the change-score outcome (mean of famine
years minus the pre-famine baseline):

```{r data-sens}
mortality$uniqueid  <- paste(as.character(mortality$provid),
                             as.character(mortality$countyid), sep = "-")
mortality$G        <- ifelse(mortality$pczupu >= median(mortality$pczupu, na.rm = TRUE), 1, 0)
mortality$lnpczupu <- log(mortality$pczupu + 1)   # log(x+1) handles the ~45% zero counties

covar <- c("avggrain", "nograin", "urban", "dis_bj", "dis_pc",
           "rice", "minority", "edu", "lnpop")

s <- fdid_prepare(
  data       = mortality,
  Y_label    = "mortality",
  X_labels   = covar,
  G_label    = "G",
  unit_label = "uniqueid",
  time_label = "year"
)

# Change-score outcome: mean(famine years) - reference year
# Covariates in s are stored as x1–x9 (matching order of X_labels above)
s$chg_mortality <- rowMeans(s[, paste0("Y_", 1958:1961)]) - s$Y_1957
xvars <- paste0("x", seq_along(covar))
```

## Binary treatment ($G$ = high vs. low social capital)

Fit the OLS change-score regression and pass it to `sensemakr`:

```{r sens-binary}
fml_bin <- as.formula(paste("chg_mortality ~ G +", paste(xvars, collapse = " + ")))
mod_bin  <- lm(fml_bin, data = s)

sens_bin <- sensemakr(
  model                = mod_bin,
  treatment            = "G",
  benchmark_covariates = xvars,
  kd                   = 1,
  sensitivity.of       = "t-value"
)
summary(sens_bin)
```

```{r plot-sens-binary, fig.width=6, fig.height=6}
plot(sens_bin, show.unadjusted = FALSE)
```

The contour plot shows the combinations of confounder-treatment and
confounder-outcome partial $R^2$ that would reduce the t-value to zero.
Dots mark benchmark values corresponding to each observed covariate.

## Continuous treatment ($\ln(\text{pczupu} + 1)$)

Repeat with `lnpczupu` as a continuous treatment:

```{r sens-continuous}
# lnpczupu = log(pczupu + 1) is included in the bundled mortality dataset
s_cont <- fdid_prepare(
  data       = transform(mortality, G = lnpczupu),
  Y_label    = "mortality",
  X_labels   = covar,
  G_label    = "G",
  unit_label = "uniqueid",
  time_label = "year"
)
s_cont$chg_mortality <- rowMeans(s_cont[, paste0("Y_", 1958:1961)]) - s_cont$Y_1957

fml_cont <- as.formula(paste("chg_mortality ~ G +", paste(xvars, collapse = " + ")))
mod_cont  <- lm(fml_cont, data = s_cont)

sens_cont <- sensemakr(
  model                = mod_cont,
  treatment            = "G",
  benchmark_covariates = xvars,
  kd                   = 1,
  sensitivity.of       = "t-value"
)
```

```{r plot-sens-continuous, fig.width=6, fig.height=6}
plot(sens_cont, show.unadjusted = FALSE)
```

## References

Cinelli, C. and Hazlett, C. (2020). "Making Sense of Sensitivity: Extending
Omitted Variable Bias." *Journal of the Royal Statistical Society: Series B*,
82(1), 39–67.

Xu, Y., Zhao, S., and Ding, P. (2026). "Factorial Difference-in-Differences."
*Journal of the American Statistical Association*.
<https://doi.org/10.1080/01621459.2026.2628343>
