# Sensitivity Analysis {#sec-sensitivity}

```{r setup-sens, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
```

After estimating treatment effects with `fdid()`, a natural next step is to
assess how robust the results are to potential unmeasured confounders. We
recommend using the [`sensemakr`](https://carloscinelli.com/sensemakr/) package
(@cinelli2020making) for this purpose.

`sensemakr` quantifies sensitivity in terms of partial $R^2$: how strongly
would an omitted variable need to be associated with both the treatment $G$ and
the change-score outcome $\Delta Y$ to explain away the estimated effect.
A full methodological discussion is in @xu2026factorial (Appendix A3).

## Setup

```{r library-sens}
library(fdid)
library(sensemakr)
data(fdid)
```

We use the prepared wide-format data from [Chapter @sec-binary] and construct
the change-score outcome: the average mortality during the famine years minus
the pre-famine baseline.

```{r data-sens}
mortality$uniqueid <- paste(as.character(mortality$provid),
                            as.character(mortality$countyid), sep = "-")
mortality$G        <- ifelse(mortality$pczupu >= median(mortality$pczupu, na.rm = TRUE), 1, 0)

covar <- c("avggrain", "nograin", "urban", "dis_bj", "dis_pc",
           "rice", "minority", "edu", "lnpop")

s <- fdid_prepare(
  data       = mortality,
  Y_label    = "mortality",
  X_labels   = covar,
  G_label    = "G",
  unit_label = "uniqueid",
  time_label = "year"
)

# Change-score outcome: mean(famine years) - reference year
# Covariates in s are stored as x1–x9 (matching order of X_labels above)
s$chg_mortality <- rowMeans(s[, paste0("Y_", 1958:1961)]) - s$Y_1957
xvars <- paste0("x", seq_along(covar))
```

## Sensitivity analysis

Fit the OLS change-score regression and pass it to `sensemakr`:

```{r sens-binary}
fml <- as.formula(paste("chg_mortality ~ G +", paste(xvars, collapse = " + ")))
mod <- lm(fml, data = s)

sens <- sensemakr(
  model                = mod,
  treatment            = "G",
  benchmark_covariates = xvars,
  kd                   = 1,
  sensitivity.of       = "t-value"
)
summary(sens)
```

```{r plot-sens, fig.width=6, fig.height=6}
plot(sens, show.unadjusted = FALSE)
```

### Interpreting the plot

The contour plot visualises how a hypothetical unmeasured confounder would need
to relate to both the treatment and the outcome to explain away the result:

- **Horizontal axis**: partial $R^2$ of the confounder with $G$ — how much
  variation in the treatment the confounder explains, beyond the observed covariates.
- **Vertical axis**: partial $R^2$ of the confounder with $\Delta Y$ — how much
  variation in the change-score outcome it explains.
- **Contour lines**: combinations that yield a given t-value. The outermost
  contour (labelled `t = 0`) marks the boundary where the effect is no longer
  statistically distinguishable from zero.
- **Dots (benchmarks)**: each dot represents one observed covariate scaled to
  have the same association strength as that covariate (`kd = 1`). If all
  benchmark dots fall well inside the t = 0 contour, the result is robust —
  an unobserved confounder would need to be substantially stronger than any
  observed covariate to nullify the finding.

In this application, the benchmarks cluster far from the t = 0 boundary,
suggesting the moderation effect of high social capital on famine mortality is
robust to moderate unmeasured confounding.

## Continuous G

Repeat the analysis with `lnpczupu` as a continuous treatment:

```{r sens-continuous}
mortality$lnpczupu <- log(mortality$pczupu + 1)

s_cont <- fdid_prepare(
  data       = transform(mortality, G = lnpczupu),
  Y_label    = "mortality",
  X_labels   = covar,
  G_label    = "G",
  unit_label = "uniqueid",
  time_label = "year"
)
s_cont$chg_mortality <- rowMeans(s_cont[, paste0("Y_", 1958:1961)]) - s_cont$Y_1957

fml_cont <- as.formula(paste("chg_mortality ~ G +", paste(xvars, collapse = " + ")))
mod_cont  <- lm(fml_cont, data = s_cont)

sens_cont <- sensemakr(
  model                = mod_cont,
  treatment            = "G",
  benchmark_covariates = xvars,
  kd                   = 1,
  sensitivity.of       = "t-value"
)
summary(sens_cont)
```

```{r plot-sens-cont, fig.width=6, fig.height=6}
plot(sens_cont, show.unadjusted = FALSE)
```

The interpretation mirrors the binary case: benchmark dots well inside the
t = 0 contour indicate that the moderation effect of social capital — measured
continuously — is robust to moderate unmeasured confounding.
