---
title: "Getting Started with fdid"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with fdid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 5
)
```

This vignette provides a minimal introduction to `fdid` using the China famine
replication data from Xu, Zhao, and Ding (2026). For a full tutorial — covering
all estimators, plotting options, and method comparisons — see the
[online Quarto book](https://yiqingxu.org/packages/fdid/).

## Installation

```{r install, eval=FALSE}
# CRAN
install.packages("fdid")

# Development version
remotes::install_github("xuyiqing/fdid")
```

## Data

The package ships with county-level mortality data from China (1954–1966). The
treatment factor `G` is a binary indicator for high social capital (above-median
genealogy book density, `pczupu`).

```{r data}
library(fdid)
data(fdid)   # loads `mortality`

# Create binary treatment and log-transform of pczupu
mortality$uniqueid <- paste(mortality$provid, mortality$countyid, sep = "-")
mortality$G        <- as.integer(mortality$pczupu >= median(mortality$pczupu, na.rm = TRUE))

head(mortality[, c("uniqueid", "year", "mortality", "G", "pczupu", "lnpczupu")])
```

## Prepare data

`fdid_prepare()` reshapes long-format panel data into the wide format required
by `fdid()`.

```{r prepare}
covariates <- c("avggrain", "nograin", "urban", "dis_bj",
                "dis_pc", "rice", "minority", "edu", "lnpop")

s <- fdid_prepare(
  data       = mortality,
  Y_label    = "mortality",
  X_labels   = covariates,
  G_label    = "G",
  unit_label = "uniqueid",
  time_label = "year"
)
```

## Estimation

`fdid()` returns pre-treatment, event-period, and post-treatment aggregates
along with dynamic (year-by-year) estimates.

```{r estimate}
result <- fdid(s, tr_period = 1958:1961, ref_period = 1957)
summary(result)
```

## Visualisation

Raw group means over time:

```{r plot-raw}
plot(result, type = "raw")
```

Dynamic (year-by-year) estimates:

```{r plot-dynamic}
plot(result, type = "dynamic")
```

Propensity score overlap (requires `method = "ipw"` or `"aipw"`):

```{r plot-overlap}
result_ipw <- fdid(s, tr_period = 1958:1961, ref_period = 1957, method = "ipw")
plot(result_ipw, type = "overlap")
```

## Sensitivity analysis

To assess robustness to omitted confounders, use the
[`sensemakr`](https://carloscinelli.com/sensemakr/) package directly. The
approach follows Xu, Zhao, and Ding (2026, Appendix A3):

```{r sensitivity, eval=requireNamespace("sensemakr", quietly = TRUE)}
library(sensemakr)

# Change-score outcome: mean(famine years) - baseline, using prepared wide data
# Covariates in s are stored as x1–x9 (matching order of X_labels above)
s$chg_mortality <- rowMeans(s[, paste0("Y_", 1958:1961)]) - s$Y_1957
xvars <- paste0("x", 1:9)   # x1=avggrain, x2=nograin, ..., x9=lnpop

fml <- as.formula(paste("chg_mortality ~ G +", paste(xvars, collapse = " + ")))
mod_sens <- lm(fml, data = s)

sens <- sensemakr(
  model                = mod_sens,
  treatment            = "G",
  benchmark_covariates = xvars,
  kd                   = 1,
  sensitivity.of       = "t-value"
)
plot(sens, show.unadjusted = FALSE)
```

Install `sensemakr` from CRAN if needed:

```{r install-sensemakr, eval=FALSE}
install.packages("sensemakr")
```

## References

Xu, Y., Zhao, S., and Ding, P. (2026). "Factorial Difference-in-Differences."
*Journal of the American Statistical Association*.
<https://doi.org/10.1080/01621459.2026.2628343>
